Plant species traits and growth rates: meta-analysis shows correlations change with plant size as predicted

### 2015-05-07 MS.tex
MS change statment of novelty of the study compared o other work by the authors Text box

This study is the first one of an upcoming series performed by the authors on the same thematic: the insertion of plant size in functional ecology. Daniel Falster will develop more deeply a growth model and tested hypotheses using expended dataset on growth and traits in one site. Emma Gray and Ian Wright are working on this thematic across all plant growth forms (tree, grass and shrub), and different environements.

@@

\textbf{Running title (36/45 characters)} Trait influence on growth along plant size

@@

\textit{Number of words in the main text (excluding abstract, acknowledgements, references, table and figure legends)} 4566/5000

\textit{Number of words in each text box} 0

@@

\textit{Number of tables} 0

\textit{Number of text boxes} 0

@@

\todo[inline, color=white] {8. list of preferred and non preferred reviewers, to list any editorial board members with whom one of the authors is in conflict: see the list \href{here}{http://onlinelibrary.wiley.com/journal/10.1111/(ISSN)1461-0248/homepage/EditorialBoard.html}}

\todo[inline, color=white] {9. To add asterisk in front of the data-source references and cites them in the text}

\todo[inline, color=white] {11. To add a clear statment in cover letter of how exactely the enclosed work is outstandingly novel relative to recent work by the author and co authors cited in the present submission}

@@

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{Iida:2014ep,Iida:2014hq,Ruger:2012jv,Falster:2011ii}.Recent theory suggests that the absence of a correlation between SLA and growth at larger sizes may be expected, when  potential influences of traits on plant function are considered \citep{Falster:2011ii, falster:2013} (see Appendix \ref{app:growth} in Supporting Informations). At larger sizes the low-cost deployment benefits of high SLA may be offset or even outweighed by the high costs of rapid leaf turnover.

@@

\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions (outlined below) from a recently developed  mechanistic
model \citep{Falster:2011ii, falster:2013}(Appendix \ref{app:growth})?
\end{enumerate}

@@

Recent theoretical work \citep{Falster:2011ii} allows hypotheses to be formulated about relationships between traits and growth. Below we provide a verbal description for each of the hypotheses; for a mathematical description see Appendix \ref{app:growth}.

@@

For WD the meta-analysis showed negative correlation with growth across all plant stages, as previously reported from individual studies \citep{Wright:2010tp,Ruger:2012jv}. It should be noted however that there are few tests of this proposition at seedling stage. In our model (Appendix \ref{app:growth}) cheaper stem construction is responsible for the predicted negative correlation. Lower xylem conductance and associated lower photosynthetic potential have also been suggested \citep[reviewed by][]{Chave:2009iy}.

For the correlation between Aarea and growth, to our knowledge no study has previously assessed this in relation to plant size. In the growth model (Appendix \ref{app:growth}), a consistent correlation irrespective of size arises from a direct effect of Aarea on plant growth rate. In juvenile stage, where most of the data were, we noted no signal of an effect of low light on the positive correlation between Aarea and growth rate.

@@

\section*{Acknowledgments}\label{Acknowledgment}

@@

\begin{Appendix} \label{sec:textbox}

\renewcommand{\thefigure}{A\arabic{figure}}
\renewcommand{\thetable}{A\arabic{table}}

\section{Derivation of hypotheses from growth equations}\label{app:growth}

@@

For the case of seedlings, eqs. \ref{eq:eq0}-\ref{eq:eq1} can be simplified considerably. Turnover of tissues and allocation to reproduction have not yet begun. It is also common also to assume all respiration rates are proportional to leaf area. Biomass growth rate then becomes a linear function of leaf area:
\begin{equation}\label{eq:eq1_seedlings}
\frac{dm_t}{dt}  \approx  a_{l} \, y \, (A_{\textrm{area}} - r_{l}).
\end{equation}
A relative growth rate in mass $(\textrm{RGR}_{\textrm{mass}})$ can be obtained by dividing both sides of this simplified eq \ref{eq:eq1_seedlings} by the total mass of living tissues $m_{t}$. The outcome is equivalent to the classical partitioning of RGR \citep{Lambers:1992bj, Cornelissen:1998ta}:
\begin{equation}\label{eq:eq1_seedlings_RGR}
\textrm{RGR}_{\textrm{mass}}  \approx \textrm{NAR} \times  \textrm{SLA} \times  \textrm{LMF},
\end{equation}
 where net assimilation rate NAR=$y \, (A_{\textrm{area}} - r)$, SLA = $a_{l}/ m_{l}$ and leaf mass fraction LMF = $m_{l}/ m_{t}$.

@@

First, the relationship between SLA and growth rate is predicted to shift from positive for seedlings to non-significant or negative in adult plants, regardless of the growth measurement used.  For seedlings (eq. \ref{eq:eq1_seedlings_RGR}) cheaper leaf construction (high SLA) should translate directly into faster relative growth rate in mass. For larger plants (eq. \ref{eq:eq1}-\ref{eq:eq4}), SLA is expected to affect growth rate positively through more cost-efficient deployment of leaf area (via $da_l/dm_t$; the same effect as in seedlings), but also negatively via increased leaf turnover (eqn 1; Table \ref{tab:trade-offs}). The effect of SLA on $da_l/dm_t$ decreases as plant size increases and increasing amounts of the plant are wood. Consequently the effect of turnover costs in slowing growth comes to dominate at larger sizes.

Second, the relationship between WD and growth rate is predicted to be negative across all plant sizes. Cheaper stem construction (low WD or stem density) also increases the marginal rate of leaf deployment per total mass increment ($da_l/dm_t$), and consequently increases growth rate in leaf area, height and diameter (eqs. \ref{eq:eq2}-\ref{eq:eq4}). Disadvantages associated with cheap stem construction take the form of higher mortality and so are not reflected in growth equations (Table \ref{tab:trade-offs}). Under this scenario, wood density is negatively correlated to growth rate across all plant stages. Because the relative expenditure on stem compared to leaf is expected to increase as plants grow larger, the relationship between WD and growth may also strengthen in larger plants.

Third, the relationship between height at maturation (Hmax) and growth rate is predicted to be positive in adults, whereas no correlation is expected for seedlings and juveniles. Hmax affects growth via the fraction of energy allocated to vegetative growth versus to seed production  ($dm_t/db$ in eq. \ref{eq:eq1}; Table \ref{tab:trade-offs}). For juveniles and seedlings this fraction is 1 for all species and so no effect is expected. But once some species begin to produce seed, then species closer to their Hmax value are expected to be allocating more to seed production and less to vegetative growth. Hence when comparing across plants at a given size, species with smaller Hmax are expected to show slower vegetative growth rates.

Fourth, the relationship between seed mass and $\textrm{RGR}_{\textrm{mass}}$ is predicted to be negative in seedlings. This is because the relative growth rate slows as plants grow larger, and large-seeded species begin larger as seedlings. For larger plants, seed mass is not expected to have any distinct influence on growth rates  (Table \ref{tab:trade-offs}).

Finally, the relationship between Aarea and growth should be positive across all plant sizes (eqn 1). The effects of benefits (higher assimilation) and costs (higher respiration) of increased Area occur within the same element of eq. \ref{eq:eq0}, and assimilation must outweigh respiration for there to be growth, so the net effect will be positive (Table \ref{tab:trade-offs}).

\newpage
\begin{Supplemnetary material}\label{sec:appendices}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

@@

\begin{table}[ht]
\centering
\caption{\linespread{2}\textbf{List of the twenty traits targeted}. Abbreviation,
definition and search term used in the meta-analysis}
\label{TableA1}
\vspace{0.5cm}
\begin{tabular}{p{3cm}p{3cm}p{8cm}}
  \hline
Trait & Definition & Search terms \\
  \hline
Aarea & Photosynthetic rate per area & "photosynthetic rate per area" OR "rate of CO2 assimilation per area" OR "Aarea" OR "Amax" OR "photosynthetic capacity per area" \\
  Amass & Photosynthetic rate per mass & "photosynthetic rate per mass" OR "rate of CO2 assimilation per mass" OR "Amass" OR "Amax" \\
  Hmax & Maximum height & "potential height" OR "maximum height" OR "Hmax" \\
  Ks & Stem specific conductance & "Ks" OR "stem specific conductance" OR "stem hydraulic conductance" OR "sapwood conductance" \\
  LA & Leaf lamina area & "leaf length" OR "leaf lamina area" OR "leaf width" OR "leaf size" \\
  LAR & Leaf area ratio & "leaf area ratio" \\
  LMR & Leaf mass ratio & "leaf mass ratio" OR "leaf mass fraction" \\
  NARarea & Net assimilation rate per area &  "net assimilation rate per area" OR "rate of dry mass increament per area" OR "NARarea" OR "NAR" \\
  Narea & Leaf nitrogen content per area & "Narea" OR "nitrogen per area" OR "N/area" OR "leaf nitrogen content per area" OR "LNCarea" OR "leaf nitrogen content" \\
  NARmass & Net assimilation rate per mass &  "net assimilation rate per mass" OR "rate of dry mass increament per mass" OR "NARmass" OR "NAR" \\
  Nmass & Leaf nitrogen content per mass & "Nmass" OR "nitrogen per mass" OR "N/mass" OR "leaf nitrogen content per mass" OR "LNCmass" OR "leaf nitrogen content" \\
  Parea & Leaf phosphorus content per area & "Parea" OR "leaf P" OR  "leaf phosphorus content per area"  OR "leaf phosphorus content" \\
  Pmass & Leaf phosphorus content per mass & "Pmass" OR "leaf P" OR  "leaf phosphorus content per mass" OR "leaf phosphorus content" \\
  SA/LA & Sapwood area per leaf area & "sapwood area per leaf area" OR "huber value" OR "leaf area per sapwood area" \\
  Seed mass & Seed mass & "seed mass" OR "seed size" OR "seed volume" \\
  SLA & Specific Leaf Area & "leaf mass per area" OR "specific leaf area" OR "leaf construction cost" \\
  Thickness & Leaf thickness & "leaf thickness"  OR "leaf tissue density" \\
  Vessel density & Vessel density & "vessel density" OR "stem conduit density" \\
  Vessel size & Vessel area or diameter & "conduit area" OR "stem conduit area" OR "vessel diameter" OR "vessel size" OR "conduit size"   \\
  WD & Wood density & "wood density" OR"WD" OR "wood specific gravity" OR "stem density" OR "stem specific density" OR "SSD" \\
   \hline
\end{tabular}
\end{table}

@@

\begin{figure}[htbp]
\centering
\includegraphics{output/FigA7d.pdf}
\end{figure}

>>>>

This study is the first one of an upcoming series of theoritical articles and empirical studies performed by the authors on the same thematic: the effect of plant size on plant strategies and functional traits.

@@

\textit{Running title (36/45 characters)} Trait influence on growth along plant size

@@

\textit{Number of words in the main text} 4583

\textit{Number of words in each text box} 750

@@

\textit{Number of tables} 1

\textit{Number of text boxes} 1

@@

@@

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{Iida:2014ep,Iida:2014hq,Ruger:2012jv,Falster:2011ii}.Recent theory suggests that the absence of a correlation between SLA and growth at larger sizes may be expected, when  potential influences of traits on plant function are considered \citep{Falster:2011ii, falster:2013} (see Text Box 1 \ref{app:growth} in Supporting Informations). At larger sizes the low-cost deployment benefits of high SLA may be offset or even outweighed by the high costs of rapid leaf turnover.

@@

\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions (outlined below) from a recently developed  mechanistic
model \citep{Falster:2011ii, falster:2013}(Text Box 1 \ref{app:growth})?
\end{enumerate}

@@

Recent theoretical work \citep{Falster:2011ii} allows hypotheses to be formulated about relationships between traits and growth. Below we provide a verbal description for each of the hypotheses; for a mathematical description see Text Box 1 \ref{app:growth}.

@@

For WD the meta-analysis showed negative correlation with growth across all plant stages, as previously reported from individual studies \citep{Wright:2010tp,Ruger:2012jv}. It should be noted however that there are few tests of this proposition at seedling stage. In our model (Text Box 1 \ref{app:growth}) cheaper stem construction is responsible for the predicted negative correlation. Lower xylem conductance and associated lower photosynthetic potential have also been suggested \citep[reviewed by][]{Chave:2009iy}.

For the correlation between Aarea and growth, to our knowledge no study has previously assessed this in relation to plant size. In the growth model (Text Box 1 \ref{app:growth}), a consistent correlation irrespective of size arises from a direct effect of Aarea on plant growth rate. In juvenile stage, where most of the data were, we noted no signal of an effect of low light on the positive correlation between Aarea and growth rate.

@@

\section*{Acknowledgements}\label{Acknowledgment}

@@

\newpage
\section*{Table} \label{sec:table}

\newpage
\section*{Text Box} \label{app:growth}

\subsection*{Derivation of hypotheses from growth equations}

@@

For the case of seedlings, eqs. \ref{eq:eq0}-\ref{eq:eq1} can be simplified considerably. Turnover of tissues and allocation to reproduction have not yet begun.

% AG: you mean turnover of other tissue than leaf?%

It is also common also to assume all respiration rates are proportional to leaf area. Biomass growth rate then becomes a linear function of leaf area:
\begin{equation}\label{eq:eq1_seedlings}
\frac{dm_t}{dt}  \approx  a_{l} \, y \, (A_{\textrm{area}} - r_{l}).
\end{equation}
A relative growth rate in mass $(\textrm{RGR}_{\textrm{mass}})$ can be obtained by dividing both sides of this simplified eq \ref{eq:eq1_seedlings} by the total mass of living tissues $m_{t}$. The outcome is equivalent to the classical partitioning of RGR \citep{Lambers:1992bj, Cornelissen:1998ta}:
\begin{equation}\label{eq:eq1_seedlings_RGR}
\textrm{RGR}_{\textrm{mass}}  \approx \textrm{NAR} \times  \textrm{SLA} \times  \textrm{LMF},
\end{equation}
 where net assimilation rate NAR=$y \, (A_{\textrm{area}} - r)$, SLA = $a_{l}/ m_{l}$ and leaf mass fraction LMF = $m_{l}/ m_{t}$.

@@

\textbf{SLA:} For seedlings (eq. \ref{eq:eq1_seedlings_RGR}) cheaper leaf construction (high SLA) should translate directly into faster relative growth rate in mass. For larger plants (eq. \ref{eq:eq1}-\ref{eq:eq4}), SLA is expected to affect growth rate positively through more cost-efficient deployment of leaf area (via $da_l/dm_t$; the same effect as in seedlings), but also negatively via increased leaf turnover (eqn 1; Table \ref{tab:trade-offs}). The effect of SLA on $da_l/dm_t$ decreases as plant size increases and increasing amounts of the plant are wood. Consequently the effect of turnover costs in slowing growth comes to dominate at larger sizes.

\textbf{WD:} Cheaper stem construction (low WD or stem density) increases the marginal rate of leaf deployment per total mass increment ($da_l/dm_t$), and consequently increases growth rate in leaf area, height and diameter (eqs. \ref{eq:eq2}-\ref{eq:eq4}). Disadvantages associated with cheap stem construction take the form of higher mortality and so are not reflected in growth equations (Table \ref{tab:trade-offs}). Under this scenario, wood density is negatively correlated to growth rate across all plant stages. Because the relative expenditure on stem compared to leaf is expected to increase as plants grow larger, the relationship between WD and growth may also strengthen in larger plants.

\textbf{Hmax:} Hmax affects growth via the fraction of energy allocated to vegetative growth versus to seed production ($dm_t/db$ in eq. \ref{eq:eq1}; Table \ref{tab:trade-offs}). For juveniles and seedlings this fraction is 1 for all species and so no effect is expected. But once some species begin to produce seed, then species closer to their Hmax value are expected to be allocating more to seed production and less to vegetative growth. Hence when comparing across plants at a given size, species with smaller Hmax are expected to show slower vegetative growth rates.

\textbf{Seed mass:} The $\textrm{RGR}_{\textrm{mass}}$ slows as plants grow larger, and large-seeded species begin larger as seedlings. For larger plants, seed mass is not expected to have any distinct influence on growth rates  (Table \ref{tab:trade-offs}).

\textbf{Aarea:} The effects of benefits (higher assimilation) and costs (higher respiration) of increased Area occur within the same element of eq. \ref{eq:eq0}, and assimilation must outweigh respiration for there to be growth, so the net effect will be positive (Table \ref{tab:trade-offs}).

\newpage
\begin{appendices}\label{sec:appendices}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

@@

\begin{table}[ht]
\centering
\caption{\textbf{List of the twenty traits targeted}. Abbreviation,
definition and search term used in the meta-analysis} 
\label{TableA1}
\vspace{0.5cm}
\begin{tabular}{p{3cm}p{3cm}p{8cm}}
  \hline
Trait & Definition & Search terms \\
  \hline
Aarea & Photosynthetic rate per area & "photosynthetic rate per area" OR "rate of CO2 assimilation per area" OR "Aarea" OR "Amax" OR "photosynthetic capacity per area" \\
  Amass & Photosynthetic rate per mass & "photosynthetic rate per mass" OR "rate of CO2 assimilation per mass" OR "Amass" OR "Amax" \\
  Hmax & Maximum height & "potential height" OR "maximum height" OR "Hmax" \\
  Ks & Stem specific conductance & "Ks" OR "stem specific conductance" OR "stem hydraulic conductance" OR "sapwood conductance" \\
  LA & Leaf lamina area & "leaf length" OR "leaf lamina area" OR "leaf width" OR "leaf size" \\
  LAR & Leaf area ratio & "leaf area ratio" \\
  LMR & Leaf mass ratio & "leaf mass ratio" OR "leaf mass fraction" \\
  NARarea & Net assimilation rate per area &  "net assimilation rate per area" OR "rate of dry mass increament per area" OR "NARarea" OR "NAR" \\
  Narea & Leaf nitrogen content per area & "Narea" OR "nitrogen per area" OR "N/area" OR "leaf nitrogen content per area" OR "LNCarea" OR "leaf nitrogen content" \\
  NARmass & Net assimilation rate per mass &  "net assimilation rate per mass" OR "rate of dry mass increament per mass" OR "NARmass" OR "NAR" \\
  Nmass & Leaf nitrogen content per mass & "Nmass" OR "nitrogen per mass" OR "N/mass" OR "leaf nitrogen content per mass" OR "LNCmass" OR "leaf nitrogen content" \\
  Parea & Leaf phosphorus content per area & "Parea" OR "leaf P" OR  "leaf phosphorus content per area"  OR "leaf phosphorus content" \\
  Pmass & Leaf phosphorus content per mass & "Pmass" OR "leaf P" OR  "leaf phosphorus content per mass" OR "leaf phosphorus content" \\
  SA/LA & Sapwood area per leaf area & "sapwood area per leaf area" OR "huber value" OR "leaf area per sapwood area" \\
  Seed mass & Seed mass & "seed mass" OR "seed size" OR "seed volume" \\
  SLA & Specific Leaf Area & "leaf mass per area" OR "specific leaf area" OR "leaf construction cost" \\
  Thickness & Leaf thickness & "leaf thickness"  OR "leaf tissue density" \\
  Vessel density & Vessel density & "vessel density" OR "stem conduit density" \\
  Vessel size & Vessel area or diameter & "conduit area" OR "stem conduit area" OR "vessel diameter" OR "vessel size" OR "conduit size"   \\
  WD & Wood density & "wood density" OR"WD" OR "wood specific gravity" OR "stem density" OR "stem specific density" OR "SSD" \\
   \hline
\end{tabular}
\end{table}

@@

\begin{figure}[htbp]
\includegraphics{output/FigA7d.pdf}
\centering
\end{figure}
